
%=============================================================================
%=============================================================================
\chapter{Preface}

During the summer before my final undergraduate year at Caltech I set
out to write a math text unlike any other, namely, one written by me.
In that respect I have succeeded beautifully.  Unfortunately, the text is
neither complete nor polished.  I have a ``Warnings and Disclaimers''
section below that is a little amusing, and an appendix on probability that 
I feel concisesly captures the essence of the subject.
However, all the material in between is in some stage of development.  
I am currently working to improve and expand this text.

This text is freely available from my web set.  Currently I'm at
\href{http://www.its.caltech.edu/~sean/}
        {http://www.its.caltech.edu/{\~{}}sean/}.
I post new versions a couple of times a year.





%=============================================================================
\section{Advice to Teachers}

If you have something worth saying, write it down.







%=============================================================================
\section{Acknowledgments}

I would like to thank Professor Saffman for advising me on
this project and the Caltech SURF program for providing
the funding for me to write the first edition of this book.




%=============================================================================
\section{Warnings and Disclaimers}

\begin{itemize}
\item This book is a work in progress.  It contains quite a few
        mistakes and typos.  I would greatly appreciate your constructive
        criticism.  You can reach me at `sean@caltech.edu'.
%
\item Reading this book impairs your ability to drive a car or operate
machinery.
%
\item This book has been found to cause drowsiness in laboratory animals.
%
\item This book contains twenty-three times the US RDA of fiber.
%
\item Caution: FLAMMABLE - Do not read while smoking or near a fire.
%
\item If infection, rash, or irritation develops, discontinue use and consult
a physician.
%
\item Warning: For external use only.  Use only as directed.  Intentional
misuse by deliberately concentrating contents can be harmful or fatal.
KEEP OUT OF REACH OF CHILDREN.
%
\item In the unlikely event of a water landing do not use this book as a
flotation device.
%
\item The material in this text is fiction; any resemblance to real theorems,
living or dead, is purely coincidental.
%
\item This is by far the most amusing section of this book.
%
\item Finding the typos and mistakes in this book is left as an exercise
for the reader.  (Eye ewes a spelling chequer from thyme too thyme,
sew their should knot bee two many
misspellings.  Though I ain't so sure the grammar's too good.)
%
\item The theorems and methods in this text are subject to change without
notice.
%
\item  This is a chain book.  If you do not make seven copies and distribute
them to your friends within ten days of obtaining this text you will suffer
great misfortune and other nastiness.
%
\item
The surgeon general has determined that excessive studying is detrimental
to your social life.
%
\item
This text has been buffered for your protection and ribbed for your
pleasure.
%
\item Stop reading this rubbish and get back to work!
\end{itemize}





%=============================================================================
\section{Suggested Use}

This text is well suited to the student, professional or lay-person.  It 
makes a superb gift.  This text has a boquet that is light and fruity, with
some earthy undertones.  It is ideal with dinner or as an apertif.
Bon apetit!



%=============================================================================
\section{About the Title}

The title is only making light of naming conventions in the sciences 
and is not an insult to engineers.   If you want to learn about some 
mathematical subject, look for books with ``Introduction'' or ``Elementary''
in the title.  If it is an ``Intermediate'' text it will be incomprehensible.
If it is ``Advanced'' then not only will it be incomprehensible, it will 
have low production qualities, i.e. a crappy typewriter font, no graphics 
and no examples.  There is an exception to this rule:  When the title also 
contains the word ``Scientists'' or ``Engineers'' the advanced book
may be quite suitable for actually learning the material.




